---
output: html_document
---
**REGRESSION MODELS PROJECT**
*Lawrence Lau*
```{r echo=FALSE}
library(datasets)
data(mtcars)
attach(mtcars)
```
**<font color=blue>Summary</font>**<br>
Given the mtcars dataset, we use a linear regression and verify the results with a t.test to see that <font color=orange>manual transmissions average `r mpg0 <- mtcars[mtcars$am=="0",]; mpg1 <- mtcars[mtcars$am=="1",]; round(mean(mpg1$mpg) - mean(mpg0$mpg), 2)` miles per gallon more than automatic transmissions</font>. We then use covariate adjustment and multiple regression models to explore other variables that could be factors, and <font color=orange>after finding a best model to hold all else equal we discover manual transmissions are more efficient by 1.49 mpg </font>.

**<font color=blue>Exploratory Data Analysis</font>**<br>
In <font color=green>Appendix 1, Boxplot - Transmission Type vs MPG</font>, the avg mpg difference between automatics (`r mpg0 <- mtcars[mtcars$am=="0",]; round(mean(mpg0$mpg), 2)`) and manuals (`r mpg1 <- mtcars[mtcars$am=="1",]; round(mean(mpg1$mpg), 2)`) is `r round(mean(mpg1$mpg), 2)-round(mean(mpg0$mpg), 2)` mpg. Let's run a hypothesis test to verify there really is a difference.
```{r echo=FALSE}
automatics <- mtcars[mtcars$am=="0",]
manuals <- mtcars[mtcars$am=="1",]
t.test(automatics$mpg, manuals$mpg)$p.value
```
With a p-value of less than .05, we reject the null hypothesis that the average miles per gallon are the same in automatics and manuals.  So there really is a difference.  But as r-squared shows, that only explains ~36% of the variance. There must be other factors.
```{r}
summary(lm(mpg~am))$r.squared
```

**<font color=blue> Models Analyses</font>**<br>
We fit each individual predictor variable against outcome MPG and rank their effect using the coefficient. 
```{r}
fit2 <- lm(mpg~., data=mtcars); fit2$coefficients
```
The greatest net effect variables, wt, carb, cyl, vs, gear, drat, and qsec, are taken and after excluding any with correlation > .5 with am, fit them into a new regression with predictor am against mpg. First, correlations. 
```{r echo=FALSE}
corr <- data.frame(cor(am,wt), cor(am,carb), cor(am,cyl), cor(am,vs), cor(am,gear), cor(am,drat), cor(am,qsec))
colnames(corr) <- c("am/wt", "am/carb", "am/cyl", "am/vs", "am/gear", "am/drat", "am/qsec")
corr
```
Next, we'll leave out any pairs with correlation > .5 and run covariate model tests.
```{r echo=FALSE}
a <- lm(mpg~am+wt); b <- lm(mpg~am+carb); c <- lm(mpg~am+cyl); d <- lm(mpg~am+vs); e <- lm(mpg~am+qsec)
fit3 <- c(a$coef, b$coef, c$coef, d$coef, e$coef)
fit3
```
We'll finally run an ANOVA test starting with predictor am and outcome mpg, then add variables individually in order of influence (most to least based on above coefficients). 
```{r}
Mod1 <- lm(mpg~am); Mod2 <- lm(mpg~am+wt); Mod3 <- lm(mpg~am+wt+vs); Mod4 <- lm(mpg~am+wt+vs+cyl)
anova <- anova(Mod1, Mod2, Mod3, Mod4)
anova[6]
```
The p-value crosses above the .05 threshold after Model 3, which means the variables in that model, am+wt+vs (transmission type, weight, and cylinder alignment) gives us our best model. A residual plot of that model (reference <font color=green>APPENDIX 2, Residual Plot, lm(mpg~am+wt+vs)</font>) confirms constant variance (no heteroskedasticity) and a normal distribution.  It's a good model fit.
```{r echo=FALSE}
lm(mpg~am+wt+vs)$coef
```
R-Squared
```{r echo=FALSE}
summary(lm(mpg~am+wt+vs))$r.squared
```
Our best model coefficient shows us that we can expect a car with a manual transmission to average 1.49 mpg better than its counterpart with an automatic transmission. The R-squared number of .8079 means this model explains ~81% of the variance-  in other words, we still have about a 19% rate of uncertainty.  

**<font color=blue>Conclusion</font>**<br>
Our initial exploratory data analysis suggested that cars with manual transmissions are more efficient by 7.24 mpg.  While a hypothesis test proved there was a difference, the r-squared number showed the transmission type was responsible for only ~36% of variance.  To find the true difference in mpg caused by transmission type, all other variables were regressed individually as a predictor variable against mpg, and the variables that had the largest net effect were selected to run a correlation test against am (transmission type). Any pairs that had a correlation > .5 (suggesting the pairs were dependent or otherwise related) were thrown out.  The remaining variables were then individually added as predictors to a model with predictor am against outcome mpg. An ANOVA test was run with those models to see which model was the best (last model for which p-value < .05). The resulting model was a regression with the predictor variables am (transmission type) + wt (weight) + vs (cylinder alignment) against outcome mpg.  The coefficients 1.49(am), -3.78(wt), and 3.62(vs) give us our best estimates that ***<font color=orange>1.)*** a standard/manual transmission is more efficient than an automatic by 1.49 mpg </font>, ***2.)*** for every 1,000 pounds heavier a car gets it loses 3.78 mpg and ***3.)*** an inline/straight cylinder shape gets 3.62 mpg more than a v-shaped arrangement.  The intercept of 30.079 tells us the average car averages 30.079 miles per gallon. These numbers explain about 81% of the variance and leave 19% uncertain.  


APPENDIX 1 - Boxplot, Transmission Type vs MPG

```{r}
par(mfrow=c(2,2))
boxplot(mpg~am, data = mtcars, col = c("tan", "tan"), xlab = "Transmission Type", ylab = "MPG", main = "Transmission Type vs MPG", xaxt="n")
axis(side=1, at=c(0,1)+1, labels=c("Automatic", "Manual"))
```

<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>
APPENDIX 2 - Residual Plot, lm(mpg~am+wt+vs)
```{r}
par(mfrow = c(2,2))
plot(lm(mpg~am+wt+vs))
```