---
title: "RegModProj"
output: html_document
---
Using the mtcars dataset, we will answer whether automatic or manual transmissions are better for MPG and quantify the difference, if any, between the two transmission types.  

In view of the limiting page requirements for this report, it's assumed the reader knows how to load the mtcars dataset and that the "am" column refers to type of transmission (0=automatic, 1=manual). The code for some of the more basic assumptions not having to do with the analyses won't be shown.

###Exploratory Data Analysis
```{r echo=FALSE}
library(datasets)
data(mtcars)
head(mtcars, 2)
```

```{r}
attach(mtcars)
par(mfrow=c(2,2), omi=c(0,0,0,0), mai=c(.7,.7,.2,.2))
plot(am, mpg, axes="FALSE", xlab="", ylab="", xaxt="n")
axis(side=1, at=c(0,1), labels=c("Automatic", "Manual"))
axis(side=2, at=seq(0,40, by=5 ))
title(xlab="Transmission Type", ylab="MPG", main="Transmission Type vs MPG", )
fit1 <- lm(mpg~am)
abline(fit1, col="purple")
box()
```
We see the data points for Automatic (0) and Manual (1) transmissions.  The mean for Automatic is `r mpg0 <- mtcars[mtcars$am=="0",]; round(mean(mpg0$mpg), 2)` mpg and the mean for Manual is `r mpg1 <- mtcars[mtcars$am=="1",]; round(mean(mpg1$mpg), 2)` mpg. It seems manual transmissions are more efficient (average higher miles per gallon) than their automatic counterparts.  But let's dig a little deeper and see if there are confounding factors.
```{r}
par(mfrow=c(2,2), omi=c(0,0,0,0), mai=c(.7,.7,.2,.2))
boxplot(mpg~am, data = mtcars, col = c("dark grey", "light grey"), xlab = "Transmission Type", ylab = "MPG", main = "Transmission Type vs MPG", xaxt="n")
axis(side=1, at=c(0,1)+1, labels=c("Automatic", "Manual"))

```

```{r}
fit2 <- lm(mpg~., data=mtcars)
```
